A tutorial on how to do the same thing more than once

### 2021-08-06 manuscript.Rmd
change formatting of software and concepts closes #197 closes #204

1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of computation**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

@@

**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
We recommend using **Git** for version control, because of its widespread adoption in the R community.

Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
Git is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3").
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit refers back to its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
Git has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
Collaborating via Git requires the repository to be uploaded somewhere.
We recommend GitHub to host Git repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.

Second, we rely on **dynamic document generation**.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
**R Markdown** uses Markdown for formatting text and R (or other programming languages) for writing the statistical analysis.
Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.

@@

To automatically resolve dependencies between project files, we rely on **dependency tracking**.
In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).

We recommend using **Make** for dependency tracking because it is language independent.
The following hypothetical example illustrates the utility of Make.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
A Makefile for this project could look like this:

@@

There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite R Markdown still makes dependencies between files unambiguous.

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

**Docker** is a popular tool for containerization.
It manages software dependencies by constructing a virtual software environment independent of the host software environment.
These so-called "Docker images" function like a virtual computer  (i.e., a sort of "sand box").
A Docker image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.

Note that the software environment of the Docker image is completely separate from software installed on your computer.
This separation is excellent for reproducibility, but takes some getting used to.
For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
Each dependency that you want to use within the Docker image must be explicitly added as a dependency.
Furthermore, using Docker mayrequire you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.

Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
While R Markdown and Git are well integrated into the R environment through RStudio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
Users do not have to have repro installed to reproduce a project;
in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown). Git tracks changes to the project over time. Make manages dependencies among the files. Docker provides a container in which the final report is built using dynamic document generation in R Markdown. Reproduced from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).  "}
# file gets downloaded in Makefile
knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
```

@@

This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to be representative of a small-scale psychological research project. 
Of course, your specific situation might involve additional, more specialized steps. 
After completing the tutorial, you should be able to customize your workflow accordingly.
The first step is to install the required software.
We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing Docker.
Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
It will assist you while you follow the tutorial.

@@

These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:

@@

The messages from `repro` try to take the user by the hand to help them solve problems.
These messages are adjusted to your specific operating system and installed dependencies.
Before you continue, we ask you to run the above commands to check Git, Make, and Docker---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
If necessary, follow any instructions presented until all checks pass.

@@

Now that your system is set up for this tutorial, we will introduce you to version control with Git.
Git tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

@@

Git tracks any files that you manually add to the repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```
Now the file `private.md` will not be committed to Git and will not be made public.

Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
When you are ready, you can begin tracking your files using Git by running:

@@

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
One way to do this is through the visual user interface in the RStudio Git pane.
Click on the empty box next to the file you want to stage.
A check mark appears to indicate that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
This stores a snapshot of the current state of the project.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and/or share the files online, you can push your local repository to a remote repository.
While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
To upload the project publically to GitHub use:

@@

R Markdown is a type of dynamic document well-suited to the RStudio user interface.
The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

One advantage of this type of markup for formatting is that it can be rendered to many different output formats---both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g. specific journal requirements.
Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
and easing the generation of citations and references to Tables and Figures.

Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
Code is separated by three backticks (and the programming language in curly braces) like here:

@@

The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.

Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
Simply put, a static document should link to the version of the code that was used to create it.
The `repro` package comes with the function `repro::current_hash()` for this purpose.
This document was created from the commit with the hash [``r repro::current_hash()``](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).

Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
These metadate contain information about the document, like the title and the output format.
Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.

@@

1. Version Control
2. Dynamic Document Creation
3. Dependency Tracking
4. Software Management

@@

Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
Some R packages add additional metadata fields.
The `repro` package adds fields to the metadata to help make the research project reproducible.
These fields are used to list all dependencies of the research project.
This includes R scripts, data files, and external packages.
The format is as follows (see everything below the line *repro:*):

@@

When you or someone else who received your project want to reproduce it, they first have to install the required software, which is Make, and Docker.
Remember, you can use the `check_*`-functions to test if these are installed:

@@

When they are setup, they can ask `repro` to explain how they should use Make and Docker to reproduce the project (or you could explain it to them):

@@

4. Let `repro` generate Docker- and Makfile:

@@

it is evaluated within the Docker container using the software within it and not the locally installed R version:

>>>>

1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
4.	*Ambiguous dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

@@

*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
We recommend using *Git* for version control, because of its widespread adoption in the R community.

`Git` tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
`Git` is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3").
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit refers back to its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In `Git`, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
`Git` has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how `Git` functions and how to use `Git` for research.
@bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
Collaborating via `Git` requires the repository to be uploaded somewhere.
We recommend GitHub to host `Git` repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.

Second, we rely on *dynamic document generation*.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
*R Markdown* uses `Markdown` for formatting text and R (or other programming languages) for writing the statistical analysis.
`Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
`R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.

@@

To automatically resolve dependencies between project files, we rely on *dependency tracking*.
In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).

We recommend using *Make* for dependency tracking because it is language independent.
The following hypothetical example illustrates the utility of `Make`.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in `R Markdown` (`manuscript.Rmd`).
A `Makefile` for this project could look like this:

@@

There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, `Make` uses a recipe to build the requirement first before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
`Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

*`Docker`* is a popular tool for containerization.
It manages software dependencies by constructing a virtual software environment independent of the host software environment.
These so-called "`Docker` images" function like a virtual computer  (i.e., a sort of "sand box").
A `Docker` image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.

Note that the software environment of the Docker image is completely separate from software installed on your computer.
This separation is excellent for reproducibility, but takes some getting used to.
For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
Each dependency that you want to use within the `Docker` image must be explicitly added as a dependency.
Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.

Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the `Dockerfile`.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), internal dependency management (using `Make`), and containerization (using `Docker`).
While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
Users do not have to have repro installed to reproduce a project;
in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (`Git`), dependency tracking (`Make`), software management (`Docker`), and dynamic document generation (`R Markdown`). `Git` tracks changes to the project over time. `Make` manages dependencies among the files. `Docker` provides a container in which the final report is built using dynamic document generation in `R Markdown`. Reproduced from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).  "}
# file gets downloaded in Makefile
knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
```

@@

This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to be representative of a small-scale psychological research project. 
Of course, your specific situation might involve additional, more specialized steps. 
After completing the tutorial, you should be able to customize your workflow accordingly.
The first step is to install the required software.
We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
It will assist you while you follow the tutorial.

@@

These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have `Git` installed:

@@

The messages from `repro` try to take the user by the hand to help them solve problems.
These messages are adjusted to your specific operating system and installed dependencies.
Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
If necessary, follow any instructions presented until all checks pass.

@@

Now that your system is set up for this tutorial, we will introduce you to version control with `Git`.
`Git` tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

@@

`Git` tracks any files that you manually add to the repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```
Now the file `private.md` will not be committed to `Git` and will not be made public.

Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
When you are ready, you can begin tracking your files using `Git` by running:

@@

For `Git` to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
One way to do this is through the visual user interface in the RStudio `Git` pane.
Click on the empty box next to the file you want to stage.
A check mark appears to indicate that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
This stores a snapshot of the current state of the project.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and/or share the files online, you can push your local repository to a remote repository.
While you can choose any `Git` service (like GitLab or BitBucket), we will use GitHub in this tutorial.
Before you upload your project to `Git`, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
To upload the project publically to GitHub use:

@@

`R Markdown` is a type of dynamic document well-suited to the RStudio user interface.
The text of an `R Markdown` is formatted by `Markdown` (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the `Markdown` syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in `Markdown`:

@@

One advantage of this type of markup for formatting is that it can be rendered to many different output formats---both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g. specific journal requirements.
Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document) which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
and easing the generation of citations and references to Tables and Figures.

Now that you are familiar with the basics of `Markdown` formatting, we turn our attention to including code into the text.
Code is separated by three backticks (and the programming language in curly braces) like here:

@@

The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
This means that, whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.

Once the `R Markdown` file has been rendered to a static document, the resulting file is decoupled from the `R Markdown` and the code that created it.
This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
To avoid ambiguity, we therefore recommend referencing the identifier of the `Git` commit at the time of rendering in the static document.
Simply put, a static document should link to the version of the code that was used to create it.
The `repro` package comes with the function `repro::current_hash()` for this purpose.
This document was created from the commit with the hash [``r repro::current_hash()``](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).

Now that you know how to write text and R code in an `R Markdown`, you need to know about metadata (also called: YAML front matter).
These metadate contain information about the document, like the title and the output format.
Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.

@@

1. *Version Control*
2. *Dynamic Document Creation*
3. *Dependency Tracking*
4. *Software Management*

@@

Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
However, `R Markdown` has been extended to hundreds of document formats which are often highly configurable by their metadata.
Some R packages add additional metadata fields.
The `repro` package adds fields to the metadata to help make the research project reproducible.
These fields are used to list all dependencies of the research project.
This includes R scripts, data files, and external packages.
The format is as follows (see everything below the line *repro:*):

@@

When you or someone else who received your project want to reproduce it, they first have to install the required software, which is `Make`, and `Docker`.
Remember, you can use the `check_*`-functions to test if these are installed:

@@

When they are setup, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):

@@

4. Let `repro` generate `Docker`- and `Makefile`:

@@

it is evaluated within the `Docker` container using the software within it and not the locally installed R version:
